---
title: "Inital soil data model summary"
author: "L Heran, K Fredrickson, M Younger, K Todd-Brown (ktoddbrown@ufl.edu)"
date: "5/1/2020"
output: 
  html_document: 
    toc: yes
bibliography: references.bib
---


```{r globalKnit, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      message=FALSE,
                      warning=FALSE)
```

```{r setup}


library(datamodelr)
library(DT)
library(plyr)
library(tidyverse)

#read in the column information
allDataModels <- read_csv(file ='../data/studyDescription.csv') %>%
  filter(!is.na(study_name), !grepl("Holmqu", study_name)) %>%
  select(study_name, table_name, column_name, long_description, unit)
  

#read in the thesorus
commonVocab <- read_csv(file ='../data/comparitiveColumnNames.csv') %>%
  select(`new name`, vanGestel, Crowther, CCRCN, ISCN3, ISRaD) %>%
  filter(!is.na(`new name`), any(!(is.na(vanGestel) | is.na(Crowther) | is.na(CCRCN) |is.na(ISCN3) | is.na(ISRaD))))
```

```{r clean}
dataModels_segments <- dlply(allDataModels, c('study_name'), function(xx){as.list(unique(sprintf('%s:%s', xx$study_name, xx$table_name)))})

all_models <- allDataModels %>%
  #filter(study_name %in% c('ISRaD', 'ISCN3', 'CCRCN', 'Crowther')) %>%
  mutate(table = sprintf('%s:%s', study_name, table_name)) %>%
  rename('column' = 'column_name') %>%
  mutate(key = ((study_name %in% c('ISRaD', 'ISCN3')) & grepl('_name$', column)) |
                        ((study_name %in% c('CCRCN')) & grepl('_id$', column)),
         ref = case_when(study_name == 'ISRaD' & grepl('^entry_name$', column) ~ 'ISRaD:metadata',
                         study_name == 'ISRaD' & grepl('^site_name$', column) ~ 'ISRaD:site',
                         study_name == 'ISRaD' & grepl('^pro_name$', column) ~ 'ISRaD:profile',
                         study_name == 'ISRaD' & grepl('^flx_name$', column) ~ 'ISRaD:flux',
                         study_name == 'ISRaD' & grepl('^lyr_name$', column) ~ 'ISRaD:layer',
                         study_name == 'ISRaD' & grepl('^ist_name$', column) ~ 'ISRaD:interstitial',
                         study_name == 'ISRaD' & grepl('^frc_name$', column) ~ 'ISRaD:fraction',
                         study_name == 'ISRaD' & grepl('^inc_name$', column) ~ 'ISRaD:incubation',
                         #ISCN3
                         study_name == 'ISCN3' & grepl('^site_name$', column) ~ 'ISCN3:site',
                         study_name == 'ISCN3' & grepl('^profile_name$', column) ~ 'ISCN3:profile',
                         study_name == 'ISCN3' & grepl('^layer_name$', column) ~ 'ISCN3:layer',
                         study_name == 'ISCN3' & grepl('^fraction_name$', column) ~ 'ISCN3:fraction',
                         #CCRCN
                         study_name == 'CCRCN' & grepl('^study_id$', column) ~ 'CCRCN:Study Information',
                         study_name == 'CCRCN' & grepl('^site_id$', column) ~ 'CCRCN:Site Level',
                         study_name == 'CCRCN' & grepl('^core_id$', column) ~ 'CCRCN:Core Level',
                         #Crowther
                         study_name == 'Crowther' & grepl('^Study$', column) ~ 'Crowther:Summary',
                         study_name == 'Crowther' & grepl('Row Labels', column) ~ 'Crowther:Summary',
                         #VanGestel
                         study_name == 'vanGestel' & grepl('site.id', column) ~ 'vanGestel:sites',
                         TRUE ~ as.character(NA)),
         ref_col = case_when(study_name == 'ISRaD' & grepl('^entry_name$', column) ~ 'entry_name',
                             study_name == 'ISRaD' & grepl('^site_name$', column) ~ 'site_name',
                         study_name == 'ISRaD' & grepl('^pro_name$', column) ~ 'pro_name',
                         study_name == 'ISRaD' & grepl('^flx_name$', column) ~ 'flx_name',
                         study_name == 'ISRaD' & grepl('^lyr_name$', column) ~ 'lyr_name',
                         study_name == 'ISRaD' & grepl('^ist_name$', column) ~ 'ist_name',
                         study_name == 'ISRaD' & grepl('^frc_name$', column) ~ 'frc_name',
                         study_name == 'ISRaD' & grepl('^inc_name$', column) ~ 'inc_name',
                         #ISCN3
                         study_name == 'ISCN3' & grepl('^study_name$', column) ~ 'study_name',
                         study_name == 'ISCN3' & grepl('^profile_name$', column) ~ 'profile_name',
                         study_name == 'ISCN3' & grepl('^layer_name$', column) ~ 'layer_name',
                         study_name == 'ISCN3' & grepl('^fraction_name$', column) ~ 'fraction_name',
                         #CCRCN
                         study_name == 'CCRCN' & grepl('^study_id$', column) ~ 'study_id',
                         study_name == 'CCRCN' & grepl('^site_id$', column) ~ 'site_id',
                         study_name == 'CCRCN' & grepl('^core_id$', column) ~ 'core_id',
                         #Crowther
                         study_name == 'Crowther' & grepl('^Study$', column) ~ 'Study',
                         study_name == 'Crowther' & grepl('Row Labels', column) ~ 'Row Labels',
                         #vanGestel
                         study_name == 'vanGestel' & grepl('site.id', column) ~ 'site.id',
                         TRUE ~ as.character(NA))) %>%
  mutate(ref = if_else(table == ref, as.character(NA), ref))
```



# Introduction and goals

The primary goal of this project was to servey the different data models used by soil carbon meta-analysis.
We looked at 5 meta-analysis that looked at some aspect of soil carbon dynamics, and in the process discovered 3 additional efforts which we intend to add to our analysis later this summer. 
Some of these studies were ongoing pre-publication (CCRCN, ISCN), ongoing with incramental publications (ISRaD:@Lawrence2020) and completed projects (Crowther:@Crowther2016, vanGestel:@vanGestel2018).

# Individual PI studies were smaller.

```{r tableCounts}
allDataModels %>% select(study_name, table_name, column_name) %>%
  unique() %>%
  group_by(study_name) %>% summarize(`Table count` = length(unique(table_name)),
                                     `Variable count` = length(unique(column_name))) %>%
  rename('Study ID' = study_name) %>%
  arrange(`Variable count`) %>%
  mutate('Multi PI?' = case_when(`Study ID` %in% c("ISCN3", "ISRaD", "CCRCN") ~ 'Yes',
                                 TRUE ~ 'No')) %>%
  knitr::kable(caption = 'Table 1: The data products varied in the number of data tables that they each contained with vanGestel only containing 4 tables and the CCRCN product containing 12. There was a wider variation in the unique variables in each study, from 40 to 351.')
```

In general individual PI projects had smaller data models (see Table 1). 
While Crowther techically had 8 data tables, most of the data was in two main data table (Figure 1).
vanGestel in contrast had 4 data tables (the smallest number in the study) organized around measurement type and site characterization (Figure 2).
In contrast, multi-PI projects had larger data tables with more complex key-ed references across them (Figure 3).
This also held true for the number of variables in each study.
The single PI studies have between 40 (Crowther) and 56 (vanGestel) unique variable names.
Multi-PI studies in contrast had between 144 (CCRCN) and 351 (ISRaD).


```{r CrowtherDM, fig.cap='Figure 1: Crawther data model.'}
temp_dm <- as.data_model(all_models %>% filter(study_name == 'Crowther'))
temp_dm2 <- dm_set_segment(temp_dm, dataModels_segments)
graph <- dm_create_graph(temp_dm2, rankdir = "RL", col_attr = c('column'))
dm_render_graph(graph)
```

```{r vanGestelDM, fig.cap='Figure 1: vanGestel data model.'}
temp_dm <- as.data_model(all_models %>% filter(study_name == 'vanGestel'))
temp_dm2 <- dm_set_segment(temp_dm, dataModels_segments)
graph <- dm_create_graph(temp_dm2, rankdir = "RL", col_attr = c('column'))
dm_render_graph(graph)
```


```{r allDM, fig.cap='Figure 3: Data models with id keys only.'}

all_dm <- as.data_model(all_models)
all_dm2 <- dm_set_segment(all_dm, dataModels_segments)

graph <- dm_create_graph(all_dm2, rankdir = "RL", col_attr = c('column'), view_type = 'keys_only')
dm_render_graph(graph)
```
# Vocabulary across studies were not obviously harmonizable.

Inital efforts to harmonize the vocabulary across studies showed over 580 unique variables out of 924 total variables across all data models.
Only 5 variables were commonly shared across all data models. 
These variables tended to focus on site location, climate, bulk density, and organic carbon precentage (see Table 2).

```{r commonVariables}
commonVocab %>%
  mutate(study_count = as.numeric(!is.na(vanGestel))+ as.numeric(!is.na(Crowther)) + 
           as.numeric(!is.na(CCRCN)) + as.numeric(!is.na(ISCN3)) + as.numeric(!is.na(ISRaD)) ) %>%
  filter(study_count > 2) %>%
  arrange(-study_count) %>% select(-study_count) %>%
  knitr::kable(caption = 'Table 2: Common variables (>2 data models) across data models.')
```

# Study feature summary

Below is a brief summary of the individual study features that were immediately apparent prior to PI interviews.

## CCRCN

  - Common features
    + location, elevation, and observation time
    + depth of core and layer
    + bulk density
    + vegitation
    + carbon (?organic) percentage and loss on ignition
    
  - Unique features
    + min/max latitude
    + detailed author information
    + 'one_liner' summary
    + break out bulk density mass/volume
    + many specific isotopes listed (Am241, C14, Cs137, Be7, Pb210, Ra226)
    + X_class is free text or control vocabulary
    + coastal specific vocabulary
      * inundation/salinity
    + anthropgenic impacts
    + core-level vs site latitude/longitude and elevation

## ISCN3

  - Common features
    + location, elevation, and observation time
    + depth of core and layer
    + bulk density
    + vegitation
    + carbon percentage and loss on ignition
    
  - Unique features
    + disturbance table
    + high level of site details
      - frost free days, ponding, runoff
    + higher then average number of layer-level info
    + fraction table only shared with ISRaD

## ISRaD

## Crowther

## vanGestel

# Next steps

All groups in this study have been contracted and most confirmed interest in particpating (communication with vanGestel is pending).
We have draft the intial questions for the long format interviews below and plan to start conducting interviews in May.
By the end of June we expect to have a more general community survey targeted more broaderly to the soil science community.

## Interview questions

  1) Why did you start this study?
  2) Describe your workflow for ingesting data sets?
  3) What decisions did you make to arrive at this workflow?
  4) How would someone get a copy of the data in this study?
  5) What would you do differently if you had to start again? What would be the same?

# References